### 2015-06-10 auspice/conclusion.tex

\section{Conclusions}

\label{sec:concl}

We presented the design, implementation, and evaluation of \auspice, a scalable, geo-distributed, key-value store. At the core of \auspice\ is a placement engine for replicating  records to achieve low lookup latency, low update cost, and high availability. We have extensively evaluated \auspice\ for a representative application: a global name service to provide name-to-address mapping for mobile devices. Our evaluation shows that \auspice's placement strategy can significantly improve the performance-cost tradeoffs struck both by commercial managed DNS services employing simplistic replication strategies today as well as previously proposed DHT-based replication alternatives with or without high mobility.