### 2015-06-10 auspice/eval-case.tex

%!TEX root = Main.tex

\subsection{End-to-end mobility case studies}
\label{sec:e2e}

Can \auspice\ serve as the basis of a complete end-to-end mobility solution? To address this question, we have developed \msocket, a user-level socket library that interoperates with \auspice, and supports all four types of endpoint mobility. %Using \auspice, \msocket\  supports connection migration, multipath communication, and mobile-to-mobile communication despite address-translating middleboxes using a distributed proxy service. 
The details of \msocket's design and implementation is the subject of a separate paper \cite{msocketTR}. Here, we {\em use} \msocket\ to show proof-of-concept of some of \auspice's capabilities.