### 2015-06-10 energy/abstract.tex

%!TEX root = New.tex
\begin{abstract}

Many datacenters are primarily used for content delivery. These datacenters are typically lightly loaded during normal operation, and hence have significant potential for saving energy. However, saving energy by shutting off servers could reduce cache-hit rates and by shutting off network components could increase datacenter network congestion; both events hurt user-experience. Due to these perceived risks, operators leave entire datacenters in an always on state foregoing all energy savings.

Towards the goal of saving energy in data center networks, we present \shrink, a cluster manager that makes power management decisions for networks and servers in a coordinated manner  to maximize energy savings, and  decides load balancing and traffic engineering  to ensure a minimal impact on user-perceived performance.
Further, \shrink\ orchestrates content transfers before server shutdown events to minimize the impact of server shutdown on cache hit rates.  We implement a prototype of \shrink\ using \TBD{TBD} library, and conduct extensive trace-driven evaluation using traces from a large CDN. Our results show that \shrink\ can reduce energy consumption by TBD-TBD\% with minimal impact on user-perceived performance and saves energy within TBD\% of the optimal strategy.

\end{abstract}