### 2015-06-10 energy/optimal.tex

\appendix
\section{Optimal Strategy}
\label{sec:optimal}

A few simplifications from the above problem statement enable us to formulate the optimal strategy as a mixed-integer program. First, we assume that the demand for a content is equally distributed  over the interval for which the optimal solution is computed. Second, the optimal strategy makes no effort at limiting the number of on-off transitions. Therefore, this optimal strategy provides at least as much energy savings, compared to a formulation which also limited the number of transitions per server (switch)  to once per day. Third, content transfers after server shutdown contribute a small fraction of overall energy, and that a heuristic for deciding content transfers is able to eliminate most of the cache-misses due to server shutdown.

%****** Why current optimal is not optimal? ******
%
%****** Depends on the utilization? ******

Why current optimal is not optimal?

(1) it does not optimize over multiple intervals?
(2) it does not decide the set of content transfers after a shutdown
(3) We are running optimal periodically, so what is the best period?
(4) its strategy to retain current placement is a heuristic scheme