### 2015-06-10 tepaper/GeantResults.tex

\documentclass[a4paper,10pt]{article}
\title{Paper Title}
\date{}
\author{}
\usepackage{geometry} 
\usepackage{amsmath}    % need for subequations
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}

\geometry{left=1.0in,right=1.0in,top=1in,bottom=1in}

\begin{document}
\section{Geant ISP Results}

MLUgraph

10 Mbps
all files

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_rates_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_rates_aggregate_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_ratios_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_ratios_aggregate_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,All day data, 10 Mbps access link}
\end{figure}

1 typical TM

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_rate_195_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_rate_195_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_ratios_195_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/10Mbps/download_ratios_195_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,a typical tm, 10 Mbps access link}
\end{figure}

No limit

all files
\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/nolimit/download_rates_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/nolimit/download_rates_aggregate_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/nolimit/download_ratios_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/nolimit/download_ratios_aggregate_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,all files ,no access link bottleneck}
\end{figure}

1 typical TM

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/nolimit/download_rate_195_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/nolimit/download_rate_195_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/nolimit/download_ratios_195_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/nolimit/download_ratios_195_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,a typical TM, no access link bottleneck}
\end{figure}

Internet bottlenecks

all files
\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/internetBW/download_rates_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/internetBW/download_rates_aggregate_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/internetBW/download_ratios_aggregate_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/internetBW/download_ratios_aggregate_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,all files internetBW}
\end{figure}

1 typical TM

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\includegraphics[scale=0.5]{images/Geant/internetBW/download_rate_206_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/internetBW/download_rate_206_5000000.pdf} \\
\includegraphics[scale=0.5]{images/Geant/internetBW/download_ratios_206_100000.pdf} &
\includegraphics[scale=0.5]{images/Geant/internetBW/download_ratios_206_5000000.pdf}
\end{tabular}
\caption{
Geant  100 KB, 5 MB files,a typical TM, internetBW}
\end{figure}

\end{document}