### 2015-07-20 conclusion.tex

\chapter{Conclusion}

This thesis focused on infrastructure service design for  a content-dominated, highly mobile Internet. We considered services for managing three types of infrastructures: an Internet service provider, a global name service and a content datacenter.

Our main contribution was to show that content placement is key to infrastructure service design and effective placement improves cost, performance, and energy-related metrics. In an ISP network, we saw that location diversity improved effective network capacity for all traffic engineering schemes, thereby blurring the difference between sub-optimal and optimal traffic engineering. In an NCDN, we saw that a simple demand-oblivious LRU caching scheme single-handedly yield network and latency costs close to the best possible even in conjunction with simple redirection and routing schemes. In Auspice, a demand-aware placement of name records resulted in significant improvents in latency and cost over existing managed DNS services, DHT-based name service designs and static placement policies. In Shrink, the effectiveness of a simple LRU caching contributed significantly to mitigating the performance impact of consolidation schemes.

Further, our work showed that, in practice, simple placement schemes perform well. Content placement problems are often NP-complete because they need to make binary decisions on whether to place content at a location or not. In practice, solving such computationally expensive problems may not be necessary because simple placement heuristics -- random placement in an ISP network, LRU caching in an NCDN and a CDC, and a simple demand-aware heuristic in Auspice -- achieve significant benefits over existing strategies and/or achieve close to optimal results as shown in this thesis.

We also studied the relative importance of optimizing placement, redirection and routing in this thesis. In an ISP network with content location diversity and a Network CDN, we showed that content placement flexibility signficantly reduces the value of a carefully engineered routing. In a global name service, a better redirection scheme does improve latencies, e.g., random-k outperforms DHT due to a better redirection scheme. However, a demand-aware placement can significantly improve cost and performance even on top of an effective redirection scheme, e.g.,  Auspice does signficantly better than random-k, which uses the same redirection scheme as Auspice, as well as managed DNS, which uses global anycast routing. Overall, these findings strengthen our thesis that placement is of key importance in infrastructure service design.
q