### 2015-07-20 shrink/conclusions.tex

%!TEX root = shrink.tex
\section{Conclusions}
Content datacenters used for storing and serving content to end-users are common today. A major barrier to  widespread adoption of server and network consolidation is \cdc\ operators' concern on the impact on SLAs or user-perceived latencies. Our work takes a step towards addressing this concern by presenting a model to quantify the energy savings vs. latency inflation trade-off in \cdc s.  A key insight, supported via experiments, is that despite server consolidation, cache hit rates remain close to an unconsolidated datacenter,  which helps mitigate the impact of consolidation on user-perceived latencies. We have designed and implemented \shrink, a system that leverages this tradeoff to yield significant energy savings while affecting user-perceived latencies in a controlled manner. \shrink's novel network-aware server consolidation algorithm reduces network energy use by up to 42\% compared to network-unaware server consolidation schemes. \shrink, in experiments based on content access traces from an Akamai datacenter, reduced energy use by 35\% compared to a baseline scheme that keeps entire datacenter always on while increasing mean latency by 8\% over it. Overall, our findings encourage deployment of consolidation techniques to reduce \cdc\ energy use.